State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-20 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
updating figure captions

%The method is appropriate for state estimation of stochastic nonlinear dynamical systems. 
\section{Results}\label{ResultsSection} The neural field model derived in Section~\ref{NeuralModelSection} was used to generate the data for state and parameter estimation. All parameters for the model are given in Table~\ref{tab:Model Parameters}. An example of the simulated field can be seen in figure~\ref{fig:experimental design}(a). This figure demonstrates the experimental design used: the sensors were placed in the centre of the field (full width at half maximum). This design allowed for uncertainty around the edges of the sensor array, which would be typical from physical data, and alleviated problems associated with boundary conditions. Figure~\ref{fig:experimental design}(b) shows example data from the sensors of the model, and figure~\ref{fig:experimental design}(c) shows the power spectral density (PSD) of the simulated time-series. The PSD shows the typical $1/f$ characteristics of intracranial EEG.

@@

@@

   	\caption{Example of experimental design and model output. \textbf{a}. Example of the neural field with sensors. The centres of the sensors are shown by crosses. \textbf{b} Example of data generated by the model through the observations. \textbf{c}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} 

@@

runs. True parameters are $\zeta=100$, $\theta_0 = 10$, $\theta_1 = -8$, $\theta_3 = 0.5$. Grey bars show distributions with wider basis functions and blue bars show improved parameter distributions with narrower basis functions.}

@@

\caption{Convergence of the parameters: averaged over 250 runs.}

@@

\subsection{Spatial Frequency Analysis} 
Using the model selection techniques described in Section~\ref{SpectralAnalysisSection}, the spatial frequency of the field was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:FieldFFT}(a) shows the spatial frequency of the field. The cut-off frequency, $\nu_c$, was taken to be \dean{??} ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of \dean{??}. This confirms our separation of \dean{??} was enough to prevent problems associated with aliasing.

@@

\begin{figure}
	\label{fig:FreqAnalysis} 
	\begin{center}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/BasisChar.pdf}}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/SensorChar.pdf}}
\subfigure[][]{\includegraphics[width=1.8in]{./Graph/ObservationFrequencyResponse.pdf}}\\
	\end{center}
	\caption{(a) Field basis function frequency response, (b) Sensor frequency response, (c) The average FFT of the observed field.}

\end{figure}
Following this, the distance between basis function, $\Delta_b$, was set to $2.5$ and the basis function width, $\sigma_{\phi}^2$, was set to 3.6, 3.16 mm full width at half maximum. \mike{Analysis said 3.9 from Parham's derivation, but 3.6 still overlaps and allows more high frequency details to be represented.}

>>>>

%The method is appropriate for state estimation of stochastic nonlinear dynamical systems. 
\section{Results}\label{ResultsSection} The neural field model derived in Section~\ref{NeuralModelSection} was used to generate the data for state and parameter estimation. All parameters for the model are given in Table~\ref{tab:Model Parameters}. An example of the simulated field can be seen in figure~\ref{fig:experimental design}(a). This figure demonstrates the experimental design used where the sensor array was placed in the centre of the field (sensor centres marked by crosses). This design allowed for uncertainty around the edges of the sensor array, which would be typical from physical data, and alleviated problems associated with boundary conditions. Figure~\ref{fig:experimental design}(b) shows example data from the sensors of the model, and figure~\ref{fig:experimental design}(c) shows the power spectral density (PSD) of the simulated time-series. The PSD shows the typical $1/f$ characteristics of intracranial EEG.
\begin{table}

\centering

@@

\caption{Parameters for neural field model used to generate data to demonstrate the estimation framework.}
\end{table}

@@

   	\caption{Example of experimental design and model output. \textbf{a}. Example of the neural field with sensors. The centres of the sensors are shown by crosses. \textbf{b} Example of data generated by the model through a subset of the observations. \textbf{c}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} 

@@

realisations. Grey bars show distributions with wider basis functions ($\sigma_{\phi} = ?$) and blue bars show improved parameter distributions with narrower basis functions ($\sigma_{\phi} = ?$). \textbf{a.} The distribution of the synaptic time constant estimates (true parameter: $\zeta=100$). \textbf{b.} The distribution of the central excitatory connectivity kernel basis function parameters estimates (true parameter: $\theta_0 = 10$). \textbf{c.} The distribution of the surround inhibition connectivity kernel basis function parameter estimates (true parameter: $\theta_1 = -8$). \textbf{d.} The distribution of the longer range excitatory connectivity kernel basis function parameter estimates (true parameter $\theta_3 = 0.5$).}

@@

\caption{Convergence of the parameters averaged over 250 runs. }

@@

\subsection{Spatial Frequency Analysis} 
Using the model selection techniques described in Section~\ref{SpectralAnalysisSection}, the spatial frequency of the field was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:FieldFFT}(a) shows the spatial frequency of the field. The cut-off frequency, $\nu_c$, was taken to be \dean{??} ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of \dean{??}. This confirms our separation of \dean{??} was enough to prevent problems associated with aliasing.

@@

\begin{figure}
	\label{fig:FreqAnalysis} 
	\begin{center}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/BasisChar.pdf}}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/SensorChar.pdf}}
\subfigure[][]{\includegraphics[width=1.8in]{./Graph/ObservationFrequencyResponse.pdf}}\\
	\end{center}
	\caption{Frequency analysis of the. \textbf{a.} The spatial frequency response of a 1-dimensional neural field basis functions project. The symmetry of the field basis functions, connectivity kernel basis functions and sensors allow for the 1-dimensional representation of the spectral characteristics. The dashed line shows the cut-off frequency (-3 dB point). \textbf{b} The spatial frequency response of a 1-dimensional model sensor. The dashed line shows the cut-off frequency (-3 dB point), indicating a wider bandwidth than the field basis function. \textbf{c.} Two 2-dimensional spatial frequency (in power dB, averaged over 200 time steps) of the observed field.} 
\end{figure}
Following this, the distance between basis function, $\Delta_{phi}$, was set to $2.5$ and the basis function width, $\sigma_{\phi}^2$, was set to 3.6, 3.16 mm full width at half maximum. \mike{Analysis said 3.9 from Parham's derivation, but 3.6 still overlaps and allows more high frequency details to be represented.}
